---
title: "Thinking about our t-test assumptions"
---

```{r, message=FALSE}
library(tidyverse)
```



Let's generate data for a two sample t-test:

```{r}
n1 = 10
n2 = 10
x = rnorm(n1)
y = rnorm(n2)
dat = tibble::tibble(
  value = c(x,y),
  group = rep(c("A","B"), times = c(n1,n2))
)
```

Visualise this:

```{r}
p = dat %>% ggplot() + 
  aes(x = group, y = value) + 
  geom_boxplot()
q = dat %>% ggplot() + 
  aes(sample = value) + 
  geom_qq_line() + geom_qq() + 
  facet_grid(~group)
q
```

Perform a t-test:

```{r}
t.test(value~group, data = dat, var.equal = TRUE)
```

Extract the p-value

```{r}
tt = t.test(value~group, data = dat, var.equal = TRUE)
tt$p.value
```

OK, let's do this a lot of times:

```{r}
B = 5000
res_p = vector("numeric", length = B)
for(i in 1:B){
  x = rnorm(n1)
  y = rnorm(n2)
  dat = tibble::tibble(
    value = c(x,y),
    group = rep(c("A","B"), times = c(n1,n2))
  )
  tt = t.test(value~group, data = dat, var.equal = TRUE)
  res_p[i] = tt$p.value
}
```

In this case we **know** the null hypothesis is true because we samples `x` and `y` from a standard normal distribution. 

How often do we see a "significant" result using a significance level of 5%?

```{r}
mean(res_p<0.05)
```

How often do we see a "significant" result using a significance level of 10%?

```{r}
mean(res_p<0.1)
```

## Normality assumption

What if we don't have normal samples?

```{r}
n1 = 10
n2 = 10
x = rchisq(n1, df = 1)
y = rchisq(n2, df = 1)
dat = tibble::tibble(
  value = c(x,y),
  group = rep(c("A","B"), times = c(n1,n2))
)
p %+% dat
q %+% dat
```


```{r}
B = 5000
res_p = vector("numeric", length = B)
for(i in 1:B){
  x = rchisq(n1, df = 1)
  y = rchisq(n2, df = 1)
  dat = tibble::tibble(
    value = c(x,y),
    group = rep(c("A","B"), times = c(n1,n2))
  )
  tt = t.test(value~group, data = dat, var.equal = TRUE)
  res_p[i] = tt$p.value
}
mean(res_p<0.1)
```

## Equal variance assumption

What about the equal variance assumption?


```{r}
B = 5000
n1 = 10
n2 = 10
res_p_welch = res_p_classic = vector("numeric", length = B)
for(i in 1:B){
  x = rnorm(n1, mean = 0, sd = 1)
  y = rnorm(n2, mean = 0, sd = 2)
  dat = tibble::tibble(
    value = c(x,y),
    group = rep(c("A","B"), times = c(n1,n2))
  )
  tt_welch = t.test(value~group, data = dat, var.equal = FALSE)
  tt_classic = t.test(value~group, data = dat, var.equal = TRUE)
  res_p_welch[i] = tt_welch$p.value
  res_p_classic[i] = tt_classic$p.value
}
mean(res_p_welch<0.05)
mean(res_p_classic<0.05)
```


