---
title: "Valuing Equity"
author: "Billy"
date: "`r Sys.Date()`"
output: html_document
---

# Equity Valuation

Similar to debt securities, equity securities are valued by calculating the PV of all cash flows generated.

These cash flows consist of:

1. Dividends
  + Payments made to shareholders
  + Future dividends are uncertain and not guaranteed
  
2. Share price
  + Market price when sold on the stock exchange

The current value of a share is influenced by its future value and the future value of future dividends


## Valuation of ordinary shares

an n year investor:

$$ E = \frac{div_{1}}{1+r_{e}} + \frac{div_{2}}{(1+r_{e})^2} + \frac{div_{2}}{(1+r_{e})^3} ... + \frac{div_{n}}{(1+r_{e})^n}$$

where:

- E = price or value of equity security
- $Div_{t}$ = dividend paid in period t
- $r_{e}$ = required return on equity
- n = number of time periods


an infinite horizon investor:

$$ E = \frac{div_{1}}{1+r_{e}} + \frac{div_{2}}{(1+r_{e})^2} + \frac{div_{2}}{(1+r_{e})^3} ... $$

Consequently, the value of a future share is given by the present value of all future dividends that the company is expected to pay:

where:

$$ E = \Sigma_{t=1}^{\infty}  \frac{Div_{t}}{(1+r_{e})^t} $$

- E = price or value of equity security
- $Div_{t}$ = dividend paid in period t
- $r_{e}$ = required return on equity


### Constant Dividend model
- assumes that companies pay a constant dividend in the long run

the value of a share can be expressed as:

$$ E = \frac{Div}{r_{e}} $$


- E = price or value of equity security
- $Div_{t}$ = dividend paid in period t
- $r_{e}$ = required return on equity


### Constant growth model
- assumes that companies pay dividends that grow at a constant rate in the long run


$$ E = \frac{Div_{1}}{r_{e} - g} $$
where:

- E = price or value of equity security
- $Div_{t}$ = dividend paid in period t
- $r_{e}$ = required return on equity
- g = constant growth rate

### Dividend discount models



$$ E = \frac{Div_{1}}{r_{e} - g} $$ $$ E = \frac{Div}{r_{e}} $$
The investors **required rate of return** is determined by
- level of interest rates in economy
- risk of the stock

the firms **growth rate of future dividensds ** is determined by
- the retention ratio = 1-payout ratio
- return on requity (ROE)



#### Estimating Growth

Estimating growth is a key issue in the application of the constant growth model

According to this mode, the dividend growth is derived from the earnings of a company
- EPS (= total earnings/number of shares outstanding) which can be paid out as dividends or retained within the company
- return on new investments is the ability to convert new investments into new earnings


Using this model, the dividend growth rate can be expressed as 

$$ g = (1 - payout ratio) \times ROE $$
where:


- g = expected rate of growth of future dividends
- 1-payout ratio = retention rate or plowback ratio
- payout ratio - dividend per share/earnings per share
- ROE = return on equity capital

Assumptions:
- payout ratio is stable
- estimated ROE is stable

example:
Our company forecasts to pay a $8.33 dividend next year, which represents 100% of its earnings. This will provide investors with a 15% expected return. Instead, we decide to plowback 40% of the earnings at the firm’s current return on equity of 25%. What is the value of the stock before and after the plowback decision?

$$ g = (1 - payout ratio) \times ROE $$



#### Two stage growth model

- it is reasonable to assume that companies cannot maintain “super-normal” growth indefinitely.
- The two-stage growth model allows for the assumption of two stages of growth – an initial phase where the growth rate is not a stable growth rate and a subsequent steady state where the growth rate is stable and expected to remain for the
long term.

#### Notes

- life expectancy is the variable that changes the outcome the least
  + its OK to assume that companies live forever
  
  
  
#Tutorial Sheet

1. You wish to value a company that paid a dividend of \$1.23 per share yesterday.  Furthermore, you  are aware that five years ago the company paid a dividend of $0.80 and in the interim years the dividend paid has grown at a constant rate. 
  + What has the annual growth rate in dividends been during the past five years? 
  
```{r, include = FALSE}
g = (1.23/0.8)^(1/5) -1
g

```
  

$0.8 \times (1 + g)^5 = 1.23$ therefore $g = (1.23/0.8)^1{1/5} -1$ = 0.0894... = 9%

  + If dividends are expected to grow at the same annual growth rate for the foreseeable future, what is the maximum price that you would be willing to pay for a share in the  company if you require a return on equity of 15% p.a.? 
  
```{r, include = FALSE}
ans = (1.23*1.09)/(0.15-0.09)
ans

```

E = $Div_{t +1}/(r - g)$

E = 22.35


