---
title: "Lecture3Formulas"
author: "Billy"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lecture 3

### Present value of multiple ammounts

The present value of a stream of cash flows can be determined using the following equation:

* PV = present value of multiple future cash flows
* $CF_{t}$ = cash flow received in period t
* r = the compound interest rate
* t = the number of periods before CF is received or paid

$$ PV = \sum_{t=1}^{n} CF_{t}/(1+r)^{t} $$

Example:

You are offered an investment that promises \$1000 in the first year, \$2000 in the second year, \$3000 in the third year and \$500 in the fourth year. If an investment opportunity of similar risk generates a return of 10% p.a. compounded monthly, what is the maximum amount that you would pay for this investment?

* The maximum amount you are willing to pay is the present value of the future cash flow

```{r}

#Method 1
rate = 1 + (0.1/12) #10% divided by 12 
PV = 1000/(1 + (0.1/12))^12 + 2000/(1 + (0.1/12))^24 + 3000/(1 + (0.1/12))^36 + 500/(1 + (0.1/12))^48
PV
  

#Method 2

EAR = (1+0.1/12)^12
PV = 1000/EAR + 2000/EAR^2 + 3000/EAR^3 + 500/EAR^4
PV


```

### Future value of multiple ammounts

The future value of a stream of cash flows can be determined by the following equation:

$$ FV = \sum_{t=1}^{n} CF_{t} (1 + r)^{n-t} $$

* FV = future value of multiple cash flows
* $CF_{t}$ = cash flow received in period t 
* r = compound interest rate
* t = the number of periods before $CF_{t}$  is received or paid

Example:
I have an opportunity to purchase an asset that will provide me with some income to supplement my regular earnings. It is expected to pay me $3150 at the end of each month. What is the accumulated value of this asset after 6 months assuming an interest rate of 5% p.a. compounding monthly?

```{r}

r = (1 + 0.05/12)

FV = (3150*(r)^(6-1) + 3150*(r)^(6-2) + 3150*(r)^(6-3) + 3150*(r)^(6-4)  + 3150*(r)^(6-5)) +3150*(r)^(6-6)
FV

```

## Annuities
Annuities are a special case of multiple cash flow streams where the cash flows are of equal size and occur at regular time intervals.

Examples include:

* A constant payment made to retirees throughout the remainder of their lives
* Rentals paid in arrears
* Interest payments on a bond




### Ordinary annuity
* Cash flows occur at the end of each period and, hence, the first cash flow occurs at the end of the first year.

To determine the present value of an **ordinary annuity** of \$A:

$$ PV = A \times \frac{1-(1+r)^{-n}} {r}  $$

* Inside the bracket is called the **annuity factor**
* PV = present value
* A = value of cash flow paid/received
* n = number of cash flows that form the annuity
* r = compound interest rate per period

Example:
If you receive $1000 at the end of each year from your grandparents for four years and reinvest the cash flows at 5% p.a. compounded annually, what is the current value of this CF stream?


```{r}
#Method 1
A = 1000
r = 0.05
n = 4

PV = A*((1-(1+r)^{-n})/r)
PV

```

To determine the future value of an **ordinary annuity** :

$$ FV = A \times \frac{(1+r)^{n}-1)}{r} $$

* PV = present value
* A = value of cash flow paid/received
* n = number of cash flows that form the annuity
* r = compound interest rate per period

NOTE:

* No cash flow at time 0
* There **IS** a cash flow at time n

Example:

If you receive $1000 at the end of each year from your grandparents for four years and reinvest the cash flows at 5% p.a. compounded annually, what is the accumulated value of this CF stream?

```{r}
A = 1000
n = 4
r = 0.05

A* ((1+r)^n - 1)/r #RHS is annuity factor


```

### Annuity due
* Cash flows occur at the beginning of each period and, hence, the first cash flow occurs immediately.


To determine the **present value** of an annuity due:

$$ PV = A + A\times \frac{1-(1+r)^{(-n-1)}}{r} $$

* if it says first payment due immediately it is an annuity due question


NOTE: Example Invalid, come back and review

Example:
What is the present value of a $1000 annuity that pays five regular semi-annual payments when the interest rate is 10% p.a. compounding quarterly with the first payment due immediately?

Step 1: find r:

Need to apply effective rate formula from class one to find r

$$ Effective Rate =  (1 + \frac{r_{nom}}{m})^{mn} - 1 $$

* $r_{nom}$ = 0.1
* m = 4
* n = 2

$$ Effective Rate =  (1 + \frac{0.1}{4})^{4\times2} - 1 $$


* A = 1000
* r = $ \frac {0.1}{4} $
* n = 5

$$ PV = 1000 + 1000\times \frac{1-(1+0.025)^{(-5-1)}}{0.025} $$


### Deferred annuity
* The first cash flow in an annuity is delayed by x periods.

To determine the **present value** of a deferred annuity:

$$ PV = A \ \times \frac{(1-(1+r)^{-n}} {r}/(1+r)^{x-1} $$
where:

* x = the number of periods before the first cash flow occurs

Example:

What is the present value of an asset that commences paying cash flows of $2 million in three years’ time for four years, when the interest rate is 8.25% p.a. compounding annually?

* r = 0.0825
* A = 2000000
* n = 4
* x = 3

$$ PV = 2000000 \ \times \frac{(1-(1+0.0825)^{-4}} {0.0825}]/(1+0.0825)^{3-1} $$

PV = $5,621,712.793

### Growing annuity
* Cash flows grow at a constant rate until the end.

$$ PV = \frac{CF_{1}}{(r -g)} \times (1 - \frac{(1+g)^{n}}{(1+r)^{n}}) $$

Example:

Your parents want to setup an investment fund to fund your 4 year Bachelors degree. The tuition fee is $40,000 per semester payable at the end of the semester. The tuition fee is expected to grow at 4% p.a. compounding semi-annually and the investment fund is predicted to return 6.5% p.a. compounding semi-annually. Calculate the amount the fund must have when you start at USYD.

* r = $\frac{6.5}{2}$
* g = $\frac{0.4}{2}$
* $CF_{1}$ = 40000
* n = 8

$$ PV = \frac{40000}{(0.0325 -0.02)} \times (1 - \frac{(1+0.02)^{8}}{(1+0.0325)^{8}}) $$

PV = 297108.06


### Equivalent annuity
* Comparison of mutually exclusive opportunities with unequal lives.

NOTE: not assessable in quizes
is assessable in exam

1. calculate present value of a cash flow stream
2. Determining an annuity whose present value is equal to this cash flow stream

Example:

You are looking at buying a new printing machine for your factory. You have two machines to choose from. Both perform the same function. Machine A lasts for 4 years whilst Machine B lasts for 7 years. The required rate of return is 10% p.a. compounding annually.


### Constant perpetuity

The present value of a constant CF perpetuity:

$$ PV = \frac{CF}{r} $$
* PV = present value of perpetuity
* CF = cash flow received/paid under the perpetuity
* r = the compound interest rate per period

Constant growth perpetuities are a special type of perpetuity where the cash flows are expected to grow at a constant rate forever.

The present value of a **constant growth perpetuity**:

$$ PV = \frac{CF_{1}}{r-g} $$

where 

* g = constant growth rate
* NOTE: r > g

Example:
If the interest rate is 8.5% p.a., which investment is most desired?

* An investment that offers you $1000 a year in perpetuity with the cash flow at the end of each year.

$$ PV_{1} = \frac{1000}{0.085} $$
$ PV_{1}  = 11764 $


* A similar investment with the $1000 receipt at the beginning of each year with the first receipt starting today.

$$ PV_{2} = 1000 + \frac{1000}{0.085} $$
$ PV_{2} = \$12764.71$

* A similar investment with the $1000 cash inflow spread evenly over each year.

Given it is spread over the year, this will automatically be less than option 2 as the "evenly spread installment" is being discounted over the year as value of cash declines

* A similar investment giving $1000 paid at the end of the first year and which
would grow at a constant rate of 0.4% p.a. forever.

$$ PV = \frac{1000}{0.085 - 0.04} $$

$PV_{4} = \$12345.68$

