---
title: "A1Data"
author: "Billy"
date: "`r Sys.Date()`"
output: html_document
---

### Downloading data

This downloads the data

```{r, warning = FALSE, message=FALSE}
library(tidyverse)
library(dplyr)
x = readr::read_tsv("DATA2x02 survey (2022) - Form responses 1 (1).tsv")

view(x)

x[-155,] #delets row with improbable data
```

the old names are bad to program with, so they have been rewritten

```{r, warning = FALSE, message = FALSE}
library(dplyr)
x = readr::read_tsv("DATA2x02 survey (2022) - Form responses 1 (1).tsv")

old_names = colnames(x)
new_names = c("timestamp","covid_positive","living_arrangements","height","uni_travel_method","uni_travel_listen","spain_budget","feel_overseas","feel_anxious","study_hrs","read_news","study_load","work","lab_zoom","social_media","gender","sleep_time","wake_time","random_number","steak_preference","dominant_hand","normal_advanced","exercise_hrs","employment_hrs","city","weekly_saving","hourly_plan","weeks_behind","assignment_on_time","used_r_before","team_role","data2x02_hrs","social_media_hrs","uni_year","sport","wam","shoe_size","decade_selection")
# overwrite the old names with the new names:
colnames(x) = new_names
# combine old and new into a data frame:
name_combo = bind_cols(New = new_names, Old = old_names)
name_combo %>% gt::gt()

```

Looking at the data thats missing may give insight into bias in variables (some people wont feel comfortable entering certain data)


```{r}
library("dplyr")
library("tidyverse")
library("data.table")
test3 <- data.table(x$uni_year, x$assignment_on_time)
test3 <- as.data.frame(test3)
test3 <- test3 %>% na.omit()
colnames(test3) = c("uni_year", "assignment_on_time")


first_always = test3 %>% select(uni_year, assignment_on_time) %>% 
  filter(uni_year =="First year" & assignment_on_time == "Always")
second_always = test3 %>% select(uni_year, assignment_on_time) %>% 
  filter(uni_year =="Second year" & assignment_on_time == "Always")
third_always = test3 %>% select(uni_year, assignment_on_time) %>% 
  filter(uni_year =="Third year" & assignment_on_time == "Always")
fourth_always = test3 %>% select(uni_year, assignment_on_time) %>% 
  filter(uni_year =="Fourth year" & assignment_on_time == "Always")
fifth_always = test3 %>% select(uni_year, assignment_on_time) %>% 
  filter(uni_year =="Fifth or higher year" & assignment_on_time == "Always")

first_sometimes = test3 %>% select(uni_year, assignment_on_time) %>% 
  filter(uni_year =="First year" & assignment_on_time == "Sometimes")
second_sometimes = test3 %>% select(uni_year, assignment_on_time) %>% 
  filter(uni_year =="Second year" & assignment_on_time == "Sometimes")
third_sometimes = test3 %>% select(uni_year, assignment_on_time) %>% 
  filter(uni_year =="Third year" & assignment_on_time == "Sometimes")
fourth_sometimes = test3 %>% select(uni_year, assignment_on_time) %>% 
  filter(uni_year =="Fourth year" & assignment_on_time == "Sometimes")
fifth_sometimes = test3 %>% select(uni_year, assignment_on_time) %>% 
  filter(uni_year =="Fifth or higher year" & assignment_on_time == "Sometimes")

first_usually = test3 %>% select(uni_year, assignment_on_time) %>% 
  filter(uni_year =="First year" & assignment_on_time == "Usually")
second_usually = test3 %>% select(uni_year, assignment_on_time) %>% 
  filter(uni_year =="Second year" & assignment_on_time == "Usually")
third_usually = test3 %>% select(uni_year, assignment_on_time) %>% 
  filter(uni_year =="Third year" & assignment_on_time == "Usually")
fourth_usually = test3 %>% select(uni_year, assignment_on_time) %>% 
  filter(uni_year =="Fourth year" & assignment_on_time == "Usually")
fifth_usually = test3 %>% select(uni_year, assignment_on_time) %>% 
  filter(uni_year =="Fifth or higher year" & assignment_on_time == "Usually")


first_never = test3 %>% select(uni_year, assignment_on_time) %>% 
  filter(uni_year =="First year" & assignment_on_time == "Never")
second_never = test3 %>% select(uni_year, assignment_on_time) %>% 
  filter(uni_year =="Second year" & assignment_on_time == "Never")
third_never = test3 %>% select(uni_year, assignment_on_time) %>% 
  filter(uni_year =="Third year" & assignment_on_time == "Never")
fourth_never = test3 %>% select(uni_year, assignment_on_time) %>% 
  filter(uni_year =="Fourth year" & assignment_on_time == "Never")
fifth_never = test3 %>% select(uni_year, assignment_on_time) %>% 
  filter(uni_year =="Fifth or higher year" & assignment_on_time == "Never")


first_year=c(nrow(first_always), nrow(first_sometimes), nrow(first_usually), nrow(first_never))
second_year =c(nrow(second_always), nrow(second_sometimes), nrow(second_usually), nrow(second_never))
third_year = c(nrow(third_always), nrow(third_sometimes), nrow(third_usually), nrow(third_never))
fourth_year = c(nrow(fourth_always), nrow(fourth_sometimes), nrow(fourth_usually), nrow(fourth_never))
fifh_year = c(nrow(fifth_always), nrow(fifth_sometimes), nrow(fifth_usually), nrow(fifth_never))

matrix_for_3 = matrix(c(nrow(first_always), nrow(first_sometimes), nrow(first_usually), nrow(first_never), 
                        nrow(second_always), nrow(second_sometimes), nrow(second_usually), nrow(second_never), 
                        nrow(third_always), nrow(third_sometimes), nrow(third_usually), nrow(third_never),
                        nrow(fourth_always), nrow(fourth_sometimes), nrow(fourth_usually), nrow(fourth_never),
                        nrow(fifth_always), nrow(fifth_sometimes), nrow(fifth_usually), nrow(fifth_never)), ncol = 5)


matrix_for_3



```

```{r}


test4 <- data.table(x$uni_year, x$sport)
test4 <- as.data.frame(test4)
test4 <- test4 %>% na.omit()
colnames(test4) = c("uni_year", "sport")


early_uni_no_sport = test4 %>% select(uni_year, sport) %>% 
  filter(sport %in% c("I don't play any sport") & uni_year %in% c("First year", "Second year"))

late_uni_no_sport = test4 %>% select(uni_year, sport) %>% 
  filter(sport %in% c("I don't play any sport") & uni_year %in% c("Third year", "Fourth year", "Fifth or higher year "))

early_uni_yes_sport = test4 %>% select(uni_year, sport) %>% 
  filter(sport != c("I don't play any sport") & uni_year %in% c("First year", "Second year"))

late_uni_yes_sport = test4 %>% select(uni_year, sport) %>% 
  filter(sport != c("I don't play any sport") & uni_year %in% c("Third year", "Fourth year", "Fifth or higher year "))

nrow(early_uni_no_sport)
nrow(early_uni_yes_sport)
matrix4 = matrix(c(nrow(early_uni_no_sport), nrow(early_uni_yes_sport), nrow(late_uni_no_sport), nrow(late_uni_yes_sport)), ncol = 2)

colnames(matrix4) = c("Early uni", "Late uni")
rownames(matrix4) = c("No sport", "Yes sport")›

```

```{r, warning = FALSE, message = FALSE}
library(visdat)
visdat::vis_miss(x)
```

```{r}
library(dplyr)
library("tidyverse")

test2 <- data.table(x$work, x$study_load)
test2 <- as.data.frame(test2)
test2 <- test2 %>% na.omit()
colnames(test2) = c("Work_load", "Study_load")


cas_full = test2 %>% select(Work_load, Study_load) %>% 
  filter(Work_load %in% c("Casual") & Study_load == "Full time")

cas_part = test2 %>% select(Work_load, Study_load) %>% 
  filter(Work_load == "Casual" & Study_load == "Part time")

full_full = test2 %>% select(Work_load, Study_load) %>% 
  filter(Work_load == "Full time" & Study_load == "Full time")

full_part = test2 %>% select(Work_load, Study_load) %>% 
  filter(Work_load == "Full time" & Study_load == "Part time")

part_full = test2 %>% select(Work_load, Study_load) %>% 
  filter(Work_load == "Part time" & Study_load == "Full time")

part_part = test2 %>% select(Work_load, Study_load) %>% 
  filter(Work_load == "Part time" & Study_load == "Part time")

dw_full = test2 %>% select(Work_load, Study_load) %>% 
  filter(Work_load == "I don't currently work" & Study_load == "Full time")

dw_part = test2 %>% select(Work_load, Study_load) %>% 
  filter(Work_load == "I don't currently work" & Study_load == "Part time")



matrix_for_test = matrix(c(nrow(cas_full), nrow(cas_part), nrow(dw_full), nrow(dw_part), 
                           nrow(full_full), nrow(full_part), nrow(part_full), nrow(part_part)), ncol = 4)
colnames(matrix_for_test) = c("Casual work", "Dont work", "Full time work", "Part time work")
rownames(matrix_for_test) = c("Full time Study", "Part time Study")


graph = matrix_for_test %>% as.data.frame() %>% 
  tibble::rownames_to_column(var = "StudyLoad") %>% 
  tidyr::pivot_longer(c("Casual", "Dont", "Full time", "Part time"), 
                      names_to = "WorkLoad", values_to = "count")


p_base = ggplot(graph, aes(x = WorkLoad, y = count, fill = StudyLoad)) + 
  theme_bw(base_size = 12) + 
  scale_fill_brewer(palette = "Set1") + 
  labs(fill = "", x = "Work Group") +
  theme(legend.position = "top")

p1 = p_base + 
  geom_bar(stat = "identity") + 
  labs(y = "Count") 
p2 = p_base + 
  geom_bar(stat = "identity", position = "fill") + 
  labs(y = "Proportion")
gridExtra::grid.arrange(p1, p2, ncol = 2)


```

```{r}
data2902students = x %>% select(wam, normal_advanced) %>% 
  filter(normal_advanced == "DATA2902")
data2902students <- data2902students %>% na.omit()

data2002students = x %>% select(wam, normal_advanced) %>% 
  filter(normal_advanced == "DATA2002")
data2002students <- data2002students %>% na.omit()


norm_adv_basictable <- table(x$normal_advanced) %>% 
                          as.data.frame() %>% 
                              setNames(c("a", "b"))


norm_adv_basictable$c = c(780,80)
norm_adv_basictable <- transform(norm_adv_basictable, r=(b/c *100))
norm_adv_basictable <- transform(norm_adv_basictable, w=c(mean(data2002students$wam), mean(data2902students$wam)))
  
norm_adv_basictable %>% gt() %>% 
  tab_header(title = "Survey Responses", subtitle = "Data2002 vs Data2902") %>% 
    cols_label(
    a = "Course",
    b = "Responses",
    c = "Students",
    r = "Response Rate (%)",
    w = "WAM"
  )
```



```{r, warning = FALSE, message = FALSE}
library(naniar)
library(ggplot2)

data_frame_of_missing <- as.data.frame(colSums(is.na(x)))

data_frame_of_missing


barplot(t(as.matrix(data_frame_of_missing)), 
        main = "Missigness of values", 
        xlab = "Number of Missing Values", 
        ylab = "Variables", 
        names.arg = c(new_names),
        col = "darkred",
        horiz = TRUE)

gg_miss_var(x) + labs(title = "Missingness of Data")

```

```{r}
library(dplyr)
library(tidyverse)
set.seed(123)

# removes hours per week that are greater than 25 and na values
data_to_test1 = x$data2x02_hrs[! x$data2x02_hrs > 25] %>% na.omit()
data_to_test1

bp1 = boxplot(data_to_test1, main = "Time Studying Data2x02", ylab = "Hours per week")



```

Cleans Height//sorts height, might be worth removing 254 cm
```{r}
x = x %>% 
  dplyr::mutate(
    height_clean = readr::parse_number(height),
    height_clean = case_when(
      height_clean <= 2.5 ~ height_clean * 100,
      height_clean <= 9 ~ NA_real_,
      TRUE ~ height_clean
    )
  )
height_clean_sorted=sort(x$height_clean)

height_clean_sorted = height_clean_sorted[- 188]
height_clean_sorted



shapiro.test(height_clean_sorted)

# Proves normal distribution
# significance above 5% means the data is valid 

```


```{r}
x = x %>% mutate(
  gender_clean = gendercoder::recode_gender(gender)
)
x %>% janitor::tabyl(
  gender, gender_clean
) %>% gt::gt() %>% 
  gt::tab_spanner(label = "Recoded outcomes", columns = 2:5) %>% 
  gt::cols_label(gender = "Original outcomes")

x$gender


```


Social media Clean
```{r}
x= x %>% mutate(
  social_media_clean = tolower(social_media),
  social_media_clean = str_replace_all(social_media_clean, '[[:punct:]]',' '),
  social_media_clean = stringr::word(social_media_clean),
  social_media_clean = case_when(
    stringr::str_starts(social_media_clean,"ins") ~ "instagram",
    stringr::str_starts(social_media_clean,"ti") ~ "tiktok",
    stringr::str_starts(social_media_clean,"mess") ~ "facebook",
    stringr::str_starts(social_media_clean,"n") ~ "none",
    is.na(social_media_clean) ~ "none",
    TRUE ~ social_media_clean
  ),
  social_media_clean = tools::toTitleCase(social_media_clean),
  social_media_clean = forcats::fct_lump_min(social_media_clean, min = 10)
)


x %>% janitor::tabyl(social_media_clean) %>% 
  arrange(desc(n)) %>% 
  gt::gt() %>% 
  gt::fmt_percent(columns = 3, decimals = 1) %>% 
  gt::cols_label(social_media_clean = "Favourite social media platform") %>% 
  gt::cols_align(align = "left", columns = 1)


```


  
SPORT CLEAN
```{r}
library(dplyr)
library(tidyverse)
x = x %>% 
  mutate(identifier = row_number()) %>% 
  mutate(sport = replace_na(sport, "I Don't Play any Sport"))

sport <- x %>%
  dplyr::select(identifier, sport) %>% 
  tidyr::separate_rows(sport,
                       sep = ", ") %>% 
  dplyr::mutate(sport = tools::toTitleCase(sport),
                sport = factor(sport),
                sport = forcats::fct_lump_min(sport, 2))






```

```{r, message = FALSE, waring=FALSE}
library(tidyverse)
library(dplyr)

sport <- x %>%
  dplyr::select(identifier, sport) %>% 
  tidyr::separate_rows(sport,
                       sep = ", ") %>% 
  dplyr::mutate(sport = tools::toTitleCase(sport),
                sport = factor(sport),
                sport = forcats::fct_lump_min(sport, 2))

sport

sport_data <- data.frame(sport)
sport_data

sports_count=dplyr::count(sport_data, sport)

sports_count
sum(sports_count$n)


aggregate(x = sports_count$n, 
          by = list(sports_count$sport),
          FUN=sum)






```




### Tinkering with data


```{r}


x %>% select(height, sport, dominant_hand) %>% filter(!complete.cases(.))

study_hrs_clean = x %>% study_hrs %>% 


```


### Practice ggplot

```{r}
library(tidyverse)
x %>% ggplot(aes(data2x02_hrs, wam)) +
  geom_point(aes(colour = study_load), alpha = 0.5) + 
  geom_smooth(method =lm) +
  labs(x = "Hours Spent Studying Data 2x02 (per week)", y = "Weighted Average Mark")
  
unique(x$data2x02_hrs)

boxplot(x$data2x02_hrs, plot = FALSE)$out


```


## practice ggstatsplot

```{r}

library("ggstatsplot")
library("rstantools")
library("ggstatsplot")
boxplot(x$data2x02_hrs, plot = FALSE)$out

boxplot(x$data2x02_hrs[!x$data2x02_hrs %in% boxplot.stats(x$data2x02_hrs)$out], ylab = "Time Studying Data2x02 (Hours per week)")


```


